---
title: "ps_4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Loading in packages 
library(tidyverse)
library(dplyr)
library(knitr)
library(readxl)
library(tidyr)
library(gt)

# Reading in the data with the appropriate column types
elections <- read_csv("ps_4_elections-poll-nc09-3.csv", col_types = cols(
  .default = col_character(),
  turnout_scale = col_double(),
  turnout_score = col_double(),
  w_LV = col_double(),
  w_RV = col_double(),
  final_weight = col_double(),
  timestamp = col_datetime(format = "")
))
```
#Question 1:

There were `r elections %>% filter(response == "Dem") %>% summarize(dems = n())` respondents who supported the Democratic candidate.

There were `r elections %>% group_by(response) %>% filter(response %in% c("Rep", "Und")) %>% summarize(favorers = n()) %>% summarize(diff = favorers[1] - favorers[2])` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (`gender` and `gender_combined`). There are `r elections %>% filter(gender != gender_combined) %>% summarize(diffs = n())` individuals for whom these variables have different values.

There are `r elections %>% filter(race_eth == "White", file_race_black != "White") %>% summarize(diffs = n())` respondents listed as “White” under `race_eth` who are not listed as “White” under `file_race_black`.

The first response of Dem came `r elections %>% filter(response %in% c("Rep", "Dem")) %>% group_by(response) %>% select(response, timestamp) %>% summarize(first = min(timestamp)) %>% summarize(diff = round(first[2] - first[1]))` minutes (rounded to the nearest minute) before the first response of Rep.

#Question 2: 

```{r echo = FALSE}
elections1 <- elections %>% 
  # Dealing only with these three variables
  select(response, race_eth, final_weight) %>%
  # Filtering out race responses (just NA) that are not in the tables
  filter(race_eth %in% c("White", "Black", "Hispanic", "Asian", "Other")) %>%
  # Graphing by response and race_eth
  group_by(response, race_eth) %>%
  # Sum the weighted values of each voter
  summarize(total = sum(final_weight)) %>%
  # Factorize to keep in order
  mutate(race_eth = factor(race_eth, levels = c("White", "Black", "Hispanic", "Asian", "Other"))) %>%
  # Spread to create Dem/Rep/Und columns
  spread(key = response, value = total)

# Replace all NAs with 0
elections1[is.na(elections1)] <- 0
# Get the percentages
elections1 <- elections1 %>%
  mutate(rowTotal = Dem + Rep + Und + `3`, Democrat = Dem/rowTotal, Republican = Rep/rowTotal, Undecided = Und/rowTotal) %>%
  # Don't want to have rowTotal or any other variabls in our table
  select(race_eth, Democrat, Republican, Undecided) %>%
  # Get rid of the column title for race_eth
  rename(" " = "race_eth")

# Table creation
gt(elections1) %>% 
  # Titles
  tab_header(
    title = md("**Party Vote Percentages By Race**"),
    subtitle = md("North Carolina’s 9th Congressional District")
  ) %>% 
  # Source
  tab_source_note(
    source_note = "Source: Siena College 2018 Live Poll Results."
  ) %>% 
  # Percentage formatting for the numerical data
  fmt_percent(columns = vars(Democrat, Republican, Undecided),
                decimals = 0) %>% 
  # Get a - for the - value
  na_if(0) %>% 
  fmt_missing(columns = vars(Undecided), rows = 4)
```

#Question 3:

```{r echo = FALSE}

elections %>%
  # Select only those who reported an education level
  filter(educ %in% c("Graduate or Professional Degree", "Bachelors' degree", "Some college or trade school", "High school", "Grade school")) %>%
  # Factorize so it stays orderly
  mutate(educ = factor(educ, levels = c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))) %>%
  # We want the data in groups
  group_by(educ) %>% 
  # Start to graph - we want a violin with a jitter over it
  ggplot(aes(x = educ, y = final_weight)) + geom_violin() +
  # Add individual points on top of violins
  geom_jitter(alpha = 0.5, size = 1, width = .25) + 
  # The way ggplot works, for some reason continuous variables have to be on the y axis
  coord_flip() +
  # Titles and captions and axes
  labs(title = "More Educated Matter Less in North Carolina 9th", subtitle = "Poll gives more weight to people who are less likely to participate in polls", y = "Weight Given to Respondent in Calculating Poll Results", x = element_blank(), caption = "New York Times Upshot/Siena College 2018 live polls")

```

#Question 4:

```{r echo = FALSE}
elections %>% 
  # Get only responses to Dem, Rep, Und, and people for which we have age data
  filter(response %in% c("Dem", "Rep", "Und"), ager != "[DO NOT READ] Refused") %>% 
  select(response, ager) %>% 
  # Group by them
  group_by(ager, response) %>% 
  # Count the number of people who responded each way
  count(response) %>%
  rename("Ages" = "ager") %>%
  # Plotting responses with color depending on age 
  ggplot(aes(x = response, y = n, fill = Ages)) + 
  # Separates out columns
  geom_col(position = "dodge2") +
  # Titles and axes
  labs(title = "Response by Age Range", caption = "Source: Siena College 2018 live polls", 
       x = "Political Party", y = "Number of Responses") 
```

#Colleagues
Chris Milne, Ilkin Bayramli
